Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-04-24 paper.tex
Two cases for teh error of rotation

Speculatively, the loss in precision should be minimal for three-point and
cluster correlation functions, because they are calculated
``in the whole volume'' of the sample. The biggest loss should be for $F_{sss}$
function because it takes into account only a small volume of the sample which
lies closely to the interface between phases and this volume is the only volume
affected by the rotation.

To estimate the error introduced by rotation, consider a
$300 \times 300 \times 300$ sample of value noise \cref{fig:roterror-noise}
which is defined as $f(x, y, z) > U$ where $f$ is a value noise function and $U$
is a threshold. We produce images of the noise rotated around $Z$ axis by $\phi$
radians by discretizing an inequation $f(p(x'), p(y'), p(z')) > U$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \phi + y \sin \phi \\
    y' &= y \cos \phi - x \sin \phi \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to
$2$ and the interface between phases is continuous. Discretizing for different
values of $\phi \in [0, \pi/4]$ we obtain the first dataset which contains the
best possible discretization of the rotated noise.

The second dataset is constructed by discretizing the noise with $\phi = 0$ and
then rotating the discretized image using nearest-neighbor interpolation.

We compute $S_3$ and $F_{sss}$ functions in periodic mode for the both datasets
and calculate pointwisely a relative error introduced by the rotation as
\begin{equation}
  \varepsilon = \frac{|F_1 - F_2|}{F_1}
  \label{eq:relerr}
\end{equation}
where $F_1$ is a correlation function for a sample from the first dataset and
$F_2$ is a correlation function for a sample from the second dataset. The we
calculate the maximal relative error for each $\phi$ as the maximum of all
pointwise errors for a given rotation angle.

Plots of maximal relative errors for functions $S_3$ and $F_{sss}$ are on
\cref{fig:roterror-error}. For the surface-surface-surface function we omit
values $F_{sss}(s_1, s_2)$ if either of its arguments is less than 10 voxels
because this function is undefined for either $s_1$ or $s_2$ being equal to zero
and our edge detection filter gives at least 7 voxel transition zone where the
correlation function is affected by this indeterminacy. The maximal error for
all $\phi$'s are 0.6\% for $S_3$ and 4.4\% for $F_{sss}$.

>>>>

Speculatively, the loss in precision should be minimal for three-point and
cluster correlation functions, because they are calculated
``in the whole volume'' of the sample. The biggest loss should be for $F_{sss}$
function because it takes into account only a small volume of the sample which
lies closely to the interface between phases and this volume is the only volume
affected by rotation.

To estimate the error introduced by rotation, consider a two-phase medium
\cref{fig:roterror-noise} whose void phase is defined by an inequality
$f(x, y, z) > U$ where $f$ is a value noise function and $U$ is a threshold. We
produce images of the noise rotated around $Z$ axis by $\phi$ radians by
discretizing an inequation $f(p(x'), p(y'), p(z')) > U$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \phi + y \sin \phi \\
    y' &= y \cos \phi - x \sin \phi \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to $2$ and
the interface between phases is continuous. Discretizing this noise for
different values of $\phi \in [0, \pi/4]$ with two different resolutions we
obtain the first dataset which contains the best possible discretization of the
rotated noise at those resolutions. For this study we have produced images with
dimensionality $300 \times 300 \times 300$ and $600 \times 600 \times 600$
voxels (i.e. the second resolution is twice as much as the first).

The second dataset is constructed by discretizing the noise with $\phi = 0$ and
then rotating the discretized image using nearest-neighbor interpolation. This
procedure is done for the same two dimensionalities as in the first dataset.

We compute $S_3$ and $F_{sss}$ functions in periodic mode for the both datasets
and calculate pointwisely a relative error introduced by rotation as
\begin{equation}
  \varepsilon = \frac{|F_1 - F_2|}{F_1}
  \label{eq:relerr}
\end{equation}
where $F_1$ is a correlation function for a sample from the first dataset and
$F_2$ is a correlation function for a sample from the second dataset for the
same combination of a rotation angle $\phi$ and dimensionality $D$. Then we
calculate the maximal relative error as the maximum of all pointwise errors for
a given pair $(\phi, D)$.

Plots of maximal relative errors for functions $S_3$ and $F_{sss}$ are on
\cref{fig:roterror-error}. When calculating the error of the
surface-surface-surface function we omit values $F_{sss}(s_1, s_2)$ if either of
its arguments is less than 10 voxels because this function is undefined for
either $s_1$ or $s_2$ being equal to zero and our edge detection filter gives at
least 7 voxel transition zone where the correlation function is affected by this
indeterminacy. The maximal error for all $\phi$'s are 0.6\% for $S_3$ and 5.0\%
for $F_{sss}$ for dimensionality $300 \times 300 \times 300$ and 0.3\% for $S_3$
and 4.5\% for $F_{sss}$ for dimensionality $600 \times 600 \times 600$.
